### 2020-03-27 Paper_EV/Paper_EV.tex
Backup update 27.03.2020

@@

@@

% Set style for pages with arabic numbering 
% This block prints only the chapter number and the chapter name, not the word "Chapter"
%\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\ \thechapter. \ #1}}{}}
% Clear current page style 
\fancyhf{}
% Head (11 pt font size and 12 pt line spacing)
\fancyhf[HRE,HLO]{\includegraphics[width=2.5cm]{figs/fsd}}
\fancyhf[HRO,HLE]{\fontsize{11}{12}\selectfont \leftmark}
% Foot (11 pt font size and 12 pt line spacing)
\fancyhf[FRE,FLO]{\fontsize{11}{12}\selectfont Design, Implem. and Eval. of an INDI Controller for a Nano-Quadrotor\\\vspace*{0.1cm}Evghenii Volodscoi}
\fancyhf[FLE,FRO]{\fontsize{11}{12}\selectfont Page \thepage \thinspace / \lastpageref{LastPages}}
% Line settings
\renewcommand{\headrulewidth}{0.1pt}	%upper line
\renewcommand{\footrulewidth}{0.1pt}	%lower line

@@

% Package for acronyms
\usepackage[acronym, automake, nopostdot, nogroupskip]{glossaries}
\setlength{\LTleft}{-0.15cm}			% left justified
\renewcommand*{\arraystretch}{1.4}		% vertical distance between entries, default is 1
% List
\newacronym{EOM}{EOM}{Equations of Motion}
\newacronym{IMU}{IMU}{Inertial Measurement Unit}
\newacronym{INDI}{INDI}{Incremental Nonlinear Dynamic Inversion}
\newacronym{MCU}{MCU}{Microcontroller}
\newacronym{NED}{NED}{North East Down}
\newacronym{NDI}{NDI}{Nonlinear Dynamic Inversion}
\newacronym{ToF}{ToF}{Time-of-Flight}
% Make gls
\renewcommand*\entryname{Acronym}
\makeglossaries

@@

\section*{Kurzfassung} \label{sec:kurzfassung}
% Start cursive block
\begin{itshape}
Deutsche Kurzfassung der Arbeit.
\end{itshape}

% Add Abstract chapter to the same page as Kurzfassung
{\let\clearpage\relax\section*{Abstract}}
% Start cursive block
\begin{itshape}
English abstract of the thesis.
\end{itshape}

@@

% Add Table of Symbols
%\chapter*{Table of Symbols} \label{sec:symbols}
\mbox{}
\nomwithdim{B}{\( F \)}{Force}{L}{$N$}
\nomwithdim{B}{\( g \)}{Gravitational acceleration}{}{$m/s^2$}
\nomwithdim{G}{\( \alpha \)}{Angle of attack}{}{$rad$}
\nomwithdim{G}{\( \zeta \)}{Damping of a linear second order system}{}{--}
\nomwithdim{S}{\( m \)}{Variable related to pitch moment}{}{}
\nomwithdim{S}{\( W \)}{Wind}{}{}
\renewcommand{\nomname}{Table of Symbols}
\printnomenclature[6em]

@@

@@

- 3. Implementation in Simulink for testing (both loops)\\
- 3. Implementation on HW (outer loop)

@@

@@

Additionally to the rigid body assumption, it is also assumed that the earth is flat and non-rotating and the reference point of the sum of all external forces acting on the body corresponds to the center of gravity of the body. Thus, the linear momentum equation is written as
\begin{equation}
	\begin{split}
		\sum (\bm{F}^G)_B = m \cdot \Big[ (\bm{\dot{v}}^G)_{B} + (\bm{\omega}^{OB})\times(\bm{v}^R)_{B} \Big]
		\label{eq:lin_momentum}
	\end{split}
\end{equation}
where $\sum(\bm{F}^G)_B \in \mathbb{R}^{3 \times 1}$ is the sum of the external forces acting on the system and applied to the center of gravity $G$, $m$ the mass of the body, $(\bm{\dot{v}}^G)_{B} \in \mathbb{R}^{3 \times 1}$ the linear acceleration of the point $G$, $(\bm{\omega}^{OB}) \in \mathbb{R}^{3 \times 1}$ the angular velocity of the body-fixed frame ($B$) with respect to the \acrfull{NED} ($O$) coordinate frame. Subscript $B$ denotes that all variables are specified in the body-fixed coordinate frame.

@@

There are different options to represent the position differential equation. Here, for completeness only, it is written as a simple relationship between the change of the position coordinates in the local \acrshort{NED} frame and the velocity coordinates of the same frame
\begin{equation}
	\begin{bmatrix}
		\dot{x}\\
		\dot{y}\\
		\dot{z}
	\end{bmatrix}_O =
	\begin{bmatrix}
		\dot{V_N}\\
		\dot{V_E}\\
		\dot{V_D}
	\end{bmatrix}_O
	\label{eq:position_diff_eq}
\end{equation}

@@

In this subsection, the \acrfull{NDI} method is explained. The \acrshort{NDI} approach is based on feedback linearization and is also called \textit{Input-Output Linearization}. Often, such type of controllers is involved in tracking control tasks, where objective is to track some desired trajectory \cite{Slotine}. To derive it, consider the following nonlinear system
\begin{subequations}
	\begin{align}
		&\bm{\dot{x}} = \bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u} \label{eq:nonlin_sys1} \\
		&\bm{y} = \bm{h}(\bm{x}) \label{eq:nonlin_sys2}
	\end{align}
	\label{eq:nonlin_sys}
\end{subequations}
\hspace{-5pt}where $\bm{x} \in \mathbb{R}^{n \times 1}$ is the state vector, $\bm{u} \in \mathbb{R}^{m \times 1}$ the input vector, $\bm{y} \in \mathbb{R}^{m \times 1}$ the output vector, $\bm{f}(\bm{x}) \in \mathbb{R}^{n \times 1}$ and $\bm{h}(\bm{x}) \in \mathbb{R}^{m \times 1}$ nonlinear vector fields and $\bm{G} \in \mathbb{R}^{m \times n}$ an input matrix. Note that the system presented in Equations (\ref{eq:nonlin_sys}) is affine in the input, which is not allways fulfilled. Using a state transformation $\bm{z} = \bm{\phi}(\bm{x})$, the affine system from Equation \ref{eq:nonlin_sys1} can be transformed into a \textit{normal} (\textit{canonical}) representation.

The core idea behind the input-output linearization method is to find a direct relationship between the desired system output and the control input. After the relationship is found it is inverted to generate the control law. To derive this relationship the output $\bm{y}$ is differentiated until the input $\bm{u}$ appears
\begin{equation}
	\begin{split}
		\bm{\dot{y}} &=  \frac{\partial\bm{y}}{\partial t} = \frac{\partial\bm{h}(\bm{x})}{\partial t} \frac{\partial\bm{x}}{\partial t} = \frac{\partial\bm{h}(\bm{x})}{\partial t} \dot{\bm{x}} = \nabla\bm{h}(\bm{x}) [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \\
		&= \nabla\bm{h}(\bm{x}) \bm{f}(\bm{x}) + \nabla\bm{h}(\bm{x}) \bm{G}(\bm{x})\bm{u} = L_{\bm{f}} \bm{h}(\bm{x}) + L_{\bm{G}} \bm{h}(\bm{x}) \bm{u} 
		\label{eq:dy}
	\end{split}
\end{equation}

@@

Now Equation (\ref{eq:dy_u_rel}) can be used to formulate the control law by solving it for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$
\begin{equation}
	\begin{split}
		\bm{u} = L_{\bm{G}}\bm{h}(\bm{x})^{-1} (\bm{\nu} - L_{\bm{f}}\bm{h}(\bm{x}))
		\label{eq:u_control_law}
	\end{split}
\end{equation}
The variable $\bm{\nu}$ is called an \textit{equivalent input} and represents the desired output of the system.

@@

The \acrshort{INDI} is an incremental form of the \acrshort{NDI} for which the lack of the accurate system dynamics model does not critically affect the performance of the control algorithm \cite{Silva}. At first the general form of the \acrshort{INDI} is presented in \ref{subsec:indi_general}, then the inner and outer control loops for the Crazyflie quadrotor are derived in \ref{subsec: indi_inner} and \ref{subsec: indi_outer}.

@@

The incremental form of the system can be obtained by taking a Taylor series expansion of the Equation \ref{eq:nonlin_sys1}
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{f}(\bm{x}_0) + \bm{G}(\bm{x}_0)\bm{u}_0 \\
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \frac{\partial}{\partial \bm{u}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor}
	\end{split}
\end{equation}
The first term on the right side of the Equation \ref{eq:nonlin_sys_taylor} is $\bm{\dot{x}_0}$. Also, evaluating the differentiation of the third term leads to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{\dot{x}}_0 \\ 
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor2}
	\end{split}
\end{equation}
The second term of the Equation \ref{eq:nonlin_sys_taylor2} contains partial derivative with respect to the state vector. Considering very small time increments of the controller loop and applying the \textit{principle of time scale separation} the second term vanishes. This is a valid assumption if the dynamics of the actuators is fast compared with the dynamics of the system \cite{Silva}. Thus, the Equation \ref{eq:nonlin_sys_taylor2} is further simplified to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} = \bm{\dot{x}}_0 + \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_indi}
	\end{split}
\end{equation}
By solving Equation \ref{eq:nonlin_sys_indi} for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$, the \acrshort{INDI} control law is obtained
\begin{equation}
	\begin{split}
		\bm{u} = \bm{u}_0 + \bm{G}(\bm{x}_0)^{-1} (\bm{\nu} - \bm{\dot{x}}_0)
		\label{eq:u_control_law_indi}
	\end{split}
\end{equation}
where $\bm{\dot{x}}_0$ is a measurable value from the previous step, $\bm{u_0}$ the control input from the previous step, $\bm{\nu}$ the reference value and $\bm{G}(\bm{x}_0)$ the control effectiveness matrix. With $\bm{\Delta u}=\bm{u}-\bm{u}_0$ the control law from Equation \ref{eq:u_control_law_indi} represents an incremental version of the Equation \ref{eq:u_control_law}. Instead of computing the complete control input command $\bm{u}$, this control law results in computing the increment of the control input $\bm{\Delta u}$ and adding it to the previous value $\bm{u}_0$. As it is less dependent on the model of the system dynamics, it is able to increase the robustness of the system \cite{Sieberling}.

\subsection{Inner loop INDI for Quadrotor} \label{subsec: indi_inner}

In this subsection, using theory from \ref{subsec:indi_general} the inner loop \acrshort{INDI} controller is derived to control angular acceleration of the Crazyflie quadrotor. The derivation of the inner loop, as well as the outer loop controllers is based on the \acrshort{INDI} controller architecture introduced by Smeur \cite{Smeur1}, \cite{Smeur2}. Equation \ref{eq:ang_momentum} from section \ref{sec:eqs_motion} serves as a basis for this derivation. The desired variable to be controlled by the inner loop \acrshort{INDI} is the angular acceleration of the quadrotor in the body-fixed coordinate frame. As in the case of a real quadrotor control problem the value of the thrust can be seen as an output of the dynamic system, it makes sense to incorporate thrust as a control variable into the control law as well. Thus, the angular momentum Equation \ref{eq:ang_momentum} is augmented with the total thrust $T$ of all four rotors \cite{Smeur2}. As only the body-fixed frame is used in the following control law derivation, the subscripts $B$ are not used in this subsection. Also all external forces and moments apply to the center of gravity of the quadrotor, thus the superscripts $G$ are also omitted. 

@@

	\end{bmatrix} \bm{\omega}

@@

As already explained in the previous subsection to derive the incremental control law the Taylor series expansion is perfromed on the Equation \ref{eq:main_eom_indi_inner}

@@

By applying differentiation and rewriting some of the terms, following equation is obtained

@@

	\bm{G}_1 = 

@@

@@

The Crazyflie 2.1 quadrotor developed by Swedish company Bitcraze belongs to the nano-quadrotor class (see Figure \ref{fig:crazyflie_drone}). Its total weight ranges from 27g to 42g, which makes possible to extend the default hardware with a pair of additional extension decks of various functionality. As it is an open source project, the main firmware, development tools and other software packages are freely available. This is one of the reasons why this small quadrotor gained a lot of popularity in different academia fields, being used for tasks reaching from the classical control theory all the way down to the obstacle avoidance problems.

The Crazyflie's main application \acrfull{MCU} is STM32F405. It runs the firmware with frequency of 1000 Hz and communicates with a large number of peripheral devices such as Bluetooth and long range radio, power management system \acrshort{MCU}, \acrfull{IMU} etc.. The \acrfull{IMU} consists of the 3-axis accelerometer, the 3-axis gyroscope and the high precision pressure sensor.

The quadrotor used in this project was extended with a flow deck sensor. The flow deck itself is equipped with two sensors: the optical flow sensor which measures the visual motion and the \acrfull{ToF} sensor which measures the range to the next object along the body-fixed z-axis of the quadcopter. By employing these sensors, the flow deck can provide estimates of the current position and linear velocity of the aircraft.

\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.4\textwidth]{figs/crazyflie_drone.jpg}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Crazyflie 2.1 nano-quadrotor]{Crazyflie 2.1 nano-quadrotor \cite{bitcraze}.}
	\label{fig:crazyflie_drone}
\end{figure}

@@

\begin{figure}[h]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=1\textwidth]{figs/sim_outer_all.pdf}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Response of the outer loop model to a step input.]{Response of the outer loop model to a step input. Blue lines represent reference values, red lines denote measured values.}
	\label{fig:sim_outer_all}
\end{figure}

@@

>>>>

% For writing angle symbol °
\usepackage{siunitx}

@@

% For SI units
\usepackage{siunitx}

@@

% Set style for pages with arabic numbering 
% This block prints only the chapter number and the chapter name, not the word "Chapter"
%\renewcommand{\chaptermark}[1]{\markboth{\MakeUppercase{\ \thechapter. \ #1}}{}}
% Clear current page style 
\fancyhf{}
% Head (11 pt font size and 12 pt line spacing)
\fancyhf[HRE,HLO]{\includegraphics[width=2.5cm]{figs/fsd}}
\fancyhf[HRO,HLE]{\fontsize{11}{12}\selectfont \leftmark}
% Foot (11 pt font size and 12 pt line spacing)
\fancyhf[FRE,FLO]{\fontsize{11}{12}\selectfont Design, Implem. and Eval. of an INDI Controller for a Nano-Quadrotor\\\vspace*{0.1cm}Evghenii Volodscoi}
\fancyhf[FLE,FRO]{\fontsize{11}{12}\selectfont Page \thepage \thinspace / \lastpageref{pagesLTS.arabic}}
% Line settings
\renewcommand{\headrulewidth}{0.1pt}	%upper line
\renewcommand{\footrulewidth}{0.1pt}	%lower line

@@

% Package for acronyms
\usepackage[acronym, automake, nopostdot, nogroupskip]{glossaries}
\setlength{\LTleft}{-0.15cm}			% left justified
\renewcommand*{\arraystretch}{1.4}		% vertical distance between entries, default is 1
% List
\newacronym{EOM}{EOM}{Equations of Motion}
\newacronym{IMU}{IMU}{Inertial Measurement Unit}
\newacronym{INDI}{INDI}{Incremental Nonlinear Dynamic Inversion}
\newacronym{MAV}{MAV}{Micro Air Vehicle}
\newacronym{MCU}{MCU}{Microcontroller}
\newacronym{NED}{NED}{North East Down}
\newacronym{NDI}{NDI}{Nonlinear Dynamic Inversion}
\newacronym{ToF}{ToF}{Time-of-Flight}
% Make gls
\renewcommand*\entryname{Acronym}
\makeglossaries

@@

\section*{Kurzfassung} \label{sec:kurzfassung}
% Start cursive block
\begin{itshape}
Inkrementelle Nichtlineare Dynamische Inversion (INDI) is ein vielversprechender Reglertyp, welcher h\"aufig f\"ur die Regelung von nichtlinearen Flugsystemen unterschiedlicher Art eingesetzt wird. Ein Vorteil der INDI ist, dass dieser Reglertyp kein detalliertes Modell der geregelten Strecke ben\"otigt. Au\ss{}erdem, wirkt INDI sehr effektiv gegen St\"orungen. Diese Semesterarbeit befasst sich mit der Auslegung eines INDI-Reglers f\"ur die Lage- und Positionsregelung eines Nano-Quadrotors. Die Arbeit beginnt mit der Herleitung des Regelalgorithmus. Der hergeleitete Regelgesetz wird dann erst in Simulink entwickelt und anschlie\ss{}end auf der Quadrotor-Hardware implementiert. Dazu werden solche Implementierungsaspekte wie Ermittlung der Eingangsparameter, Messung und Berechnung der Aktuatorzeitkonstante sowie Ermittlung des Schub-Mapping-Parameters besprochen. Abschlie\ss{}end, wird der implementierte Regler auf die F\"ahigkeit St\"orungen zu bek\"ampfen getestet.
\end{itshape}

% Add Abstract chapter to the same page as Kurzfassung
{\let\clearpage\relax\section*{Abstract}}
% Start cursive block
\begin{itshape}
Incremental Nonlinear Dynamic Inversion (INDI) is a promising control technique, widely used for control of different types of aircraft systems. Besides providing high-performance nonlinear control, this controller type does not require a detailed model of the controlled aircraft and is effective against disturbances. This semester thesis describes the development of an INDI controller to control the attitude and the position of a nano-quadrotor. It begins with the derivation of the control algorithm. The controller is then firstly developed in the Simulink environment and afterwards implemented on the quadrotor hardware.  Subsequently, the implementation aspects of the INDI controller such as estimation of the control effectiveness, measurement of the actuator time constant and estimation of the thrust mapping parameter are discussed. Finally, the implemented controller is tested on the ability to cope with disturbances.  
\end{itshape}

@@

% Add Table of Symbols
%\chapter*{Table of Symbols} \label{sec:symbols}
\mbox{}
\nomwithdim{B}{\( \bm{F} \)}{Force}{L}{$N$}
\nomwithdim{B}{\( g \)}{Gravitational acceleration}{}{$m/s^2$}
\nomwithdim{G}{\( \alpha \)}{Angle of attack}{}{$rad$}
\nomwithdim{G}{\( \zeta \)}{Damping of a linear second order system}{}{--}
\nomwithdim{S}{\( m \)}{Variable related to pitch moment}{}{}
\nomwithdim{S}{\( W \)}{Wind}{}{}
\renewcommand{\nomname}{Table of Symbols}
\printnomenclature[6em]

@@

The popularity of Micro Air Vehicles (MAVs) in research and industry has grown over the past years. The large interest in \acrshort{MAV}s can be expalined with their potential of performing a wide range of civilian and military tasks, reaching from the investigation of hazardous regions up to the agriculture mapping \cite{Ward}. One of the most popular representatives of \acrshort{MAV}s are quadrotors, belonging to the rotary-wing \acrshort{MAV}s. Their dynamic behaviour is usually of a nonlinear character. Thus, to control such an aircraft, robust control strategies have to be applied. One such strategy is called \acrfull{INDI}, which is the incremental form of the nonlinearity cancelling technique \acrfull{NDI}. The \acrshort{INDI} technique makes it possible to control an aircraft without fully modelling its dynamics. Instead, \acrshort{INDI} uses a primitive aircraft model, which eases the development and implementation of the control algorithm. In addition, \acrshort{INDI} has good properties of coping with unmodelled gust and disturbances.

@@

The main goal of this thesis was to implement an \acrshort{INDI} controller to control the linear acceleration of a nano-quadrotor. Furthermore, to have the ability to control the quadrotors position, the \acrshort{INDI} loop had to be augmented with a velocity as well as a position controller. For the role of the research quadrotor, a small quadrotor named Crazyflie 2.1 was employed. Besides the derivation of the control algorithm itself, this thesis covers the process of estimation of the control effectiveness terms, as well as the estimation of other parameters such as actuator dynamics time constant and thrust mapping parameter, which are necessary for proper functioning of the controller. Additionally to the hardware implementation of the control algorithm, this work describes the development of the controller as well as the quadrotor model in Simulink. Finally, it discusses the capability of the successfully implemented \acrshort{INDI} controller to cope with disturbances.

%- Implementation in Simulink for testing (both loops)\\
%- Implementation on HW (outer loop)  \\

@@

This semester thesis is organized into five chapters. The introductory chapter is followed by the second chapter, containing the theoretical background. Here, after introducing dynamic equations of motion of an aircraft (\ref{sec:eqs_motion}), \acrshort{NDI} (\ref{sec:ndi}) and \acrshort{INDI} (\ref{sec:indi}) control algorithms are discussed. The \acrshort{INDI} technique is then used to derive the attitude (\ref{subsec: indi_inner}) and position (\ref{subsec: indi_outer}) controllers for the Crazyflie quadrotor. Chapter three focuses on the implementation aspects of the \acrshort{INDI} algorithm derived in the previous chapter. It starts with a short description of the quadrotor hardware (\ref{sec:research_quadrotor}). In the second part (\ref{sec:simulink_model}) of the third chapter, simplified model of the Crazyflie quadrotor as well as the attitude and position controllers are developed in the Simulink environment. The third section (\ref{sec:implementation_on_hardware}) of the same chapter covers aspects that are relevant for the hardware implementation of the \acrshort{INDI} control algorithm, such as the estimation of the control effectiveness terms and thrust mapping parameter. After controller was implemented, it was tested and validated. The test results are summarized in the fourth chapter. The final chapter contains a short conclusion of the performed work.

@@

Additionally to the rigid body assumption, it is also assumed that the earth is flat and non-rotating and the reference point of the sum of all external forces acting on the body corresponds to the center of gravity of the body. Thus, the linear momentum equation is written as
\begin{equation}
	\begin{split}
		\sum (\bm{F}^G)_B = m \cdot \Big[ (\bm{\dot{v}}^G)_{B} + (\bm{\omega}^{OB})\times(\bm{v}^R)_{B} \Big]
		\label{eq:lin_momentum}
	\end{split}
\end{equation}
where $\sum(\bm{F}^G)_B \in \mathbb{R}^{3 \times 1}$ is the sum of the external forces acting on the system and applied to the center of gravity $G$, $m$ the mass of the body, $(\bm{\dot{v}}^G)_{B} \in \mathbb{R}^{3 \times 1}$ the linear acceleration of the point $G$, $(\bm{v}^R)_{B} \in \mathbb{R}^{3 \times 1}$ the velocity of any reference point $R$, $(\bm{\omega}^{OB}) \in \mathbb{R}^{3 \times 1}$ the angular velocity of the body-fixed frame ($B$) with respect to the \acrfull{NED} ($O$) coordinate frame. Subscript $B$ denotes that all variables are specified in the body-fixed coordinate frame.

@@

There are different options to represent the position differential equation. Here, for completeness only, it is written as a simple relationship between the change of the position coordinates in the local \acrshort{NED} frame and the velocity coordinates of the same frame
\begin{equation}
	\begin{bmatrix}
		\dot{x}\\
		\dot{y}\\
		\dot{z}
	\end{bmatrix}_O =
	\begin{bmatrix}
		V_N\\
		V_E\\
		V_D
	\end{bmatrix}_O
	\label{eq:position_diff_eq}
\end{equation}

@@

In this subsection, the \acrfull{NDI} method is explained. The \acrshort{NDI} approach is based on feedback linearization and is also called \textit{Input-Output Linearization}. Often, such type of controllers is involved in tracking control tasks, where the objective is to track some desired trajectory \cite{Slotine}. To derive it, consider the following nonlinear system
\begin{subequations}
	\begin{align}
		&\bm{\dot{x}} = \bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u} \label{eq:nonlin_sys1} \\
		&\bm{y} = \bm{h}(\bm{x}) \label{eq:nonlin_sys2}
	\end{align}
	\label{eq:nonlin_sys}
\end{subequations}
\hspace{-5pt}where $\bm{x} \in \mathbb{R}^{n \times 1}$ is the state vector, $\bm{u} \in \mathbb{R}^{m \times 1}$ the input vector, $\bm{y} \in \mathbb{R}^{m \times 1}$ the output vector, $\bm{f}(\bm{x}) \in \mathbb{R}^{n \times 1}$ and $\bm{h}(\bm{x}) \in \mathbb{R}^{m \times 1}$ nonlinear vector fields and $\bm{G} \in \mathbb{R}^{m \times n}$ an input matrix. Note that the system presented in Equations (\ref{eq:nonlin_sys}) is affine in the input, which is not allways fulfilled. Using a state transformation $\bm{z} = \bm{\phi}(\bm{x})$, the affine system from Equation \ref{eq:nonlin_sys1} can be transformed into a \textit{normal} (\textit{canonical}) representation.

The core idea behind the input-output linearization method is to find a direct relationship between the desired system output and the control input. After the relationship is found it is inverted to generate the control law. To derive this relationship the output $\bm{y}$ is differentiated until the input $\bm{u}$ appears
\begin{equation}
	\begin{split}
		\bm{\dot{y}} &=  \frac{\partial\bm{y}}{\partial t} = \frac{\partial\bm{h}(\bm{x})}{\partial x} \frac{\partial\bm{x}}{\partial t} = \frac{\partial\bm{h}(\bm{x})}{\partial x} \dot{\bm{x}} = \nabla\bm{h}(\bm{x}) [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \\
		&= \nabla\bm{h}(\bm{x}) \bm{f}(\bm{x}) + \nabla\bm{h}(\bm{x}) \bm{G}(\bm{x})\bm{u} = L_{\bm{f}} \bm{h}(\bm{x}) + L_{\bm{G}} \bm{h}(\bm{x}) \bm{u} 
		\label{eq:dy}
	\end{split}
\end{equation}

@@

Now Equation (\ref{eq:dy_u_rel}) can be used to formulate the control law by solving it for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$
\begin{equation}
	\begin{split}
		\bm{u} = (L_{\bm{G}}\bm{h}(\bm{x}))^{-1} (\bm{\nu} - L_{\bm{f}}\bm{h}(\bm{x}))
		\label{eq:u_control_law}
	\end{split}
\end{equation}
The variable $\bm{\nu}$ is called an \textit{virtual control input} and represents the desired output of the system.

@@

\acrshort{INDI} is an incremental form of \acrshort{NDI} for which the lack of the accurate system dynamics model does not critically affect the performance of the control algorithm \cite{Silva}. At first the general form of the \acrshort{INDI} is presented in \ref{subsec:indi_general}, then the inner and outer control loops for the Crazyflie quadrotor are derived in \ref{subsec: indi_inner} and \ref{subsec: indi_outer}.

@@

Assuming that the output variable corresponds to the state varible ($y=x$), the incremental form of the system can be obtained by taking a Taylor series expansion of the Equation \ref{eq:nonlin_sys1}
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{f}(\bm{x}_0) + \bm{G}(\bm{x}_0)\bm{u}_0 \\
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \frac{\partial}{\partial \bm{u}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor}
	\end{split}
\end{equation}
The first term on the right side of the Equation \ref{eq:nonlin_sys_taylor} is $\bm{\dot{x}_0}$. Also, evaluating the differentiation of the third term leads to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} &= \bm{\dot{x}}_0 \\ 
		&+ \frac{\partial}{\partial \bm{x}} [\bm{f}(\bm{x}) + \bm{G}(\bm{x})\bm{u}] \bigg|_{\bm{x}=\bm{x}_0,\bm{u}=\bm{u}_0} (\bm{x}-\bm{x}_0) \\
		&+ \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_taylor2}
	\end{split}
\end{equation}
The second term of the Equation \ref{eq:nonlin_sys_taylor2} contains partial derivative with respect to the state vector. Considering very small time increments of the controller loop and applying the \textit{principle of time scale separation} the second term vanishes. This is a valid assumption if the dynamics of the actuators are fast compared with the dynamics of the system \cite{Silva}. Thus, Equation \ref{eq:nonlin_sys_taylor2} is further simplified to 
\begin{equation}
	\begin{split}
		\bm{\dot{x}} = \bm{\dot{x}}_0 + \bm{G}(\bm{x}_0) (\bm{u}-\bm{u}_0) 
		\label{eq:nonlin_sys_indi}
	\end{split}
\end{equation}
By solving Equation \ref{eq:nonlin_sys_indi} for $\bm{u}$ and substituting $\bm{\dot{y}}$ with $\bm{\nu}$, the \acrshort{INDI} control law is obtained
\begin{equation}
	\begin{split}
		\bm{u} = \bm{u}_0 + \bm{G}(\bm{x}_0)^{-1} (\bm{\nu} - \bm{\dot{x}}_0)
		\label{eq:u_control_law_indi}
	\end{split}
\end{equation}
where $\bm{\dot{x}}_0$ is a measurable value from the previous step, $\bm{u_0}$ the control input from the previous step, $\bm{\nu}$ the reference value and $\bm{G}(\bm{x}_0)$ the control effectiveness matrix. With $\bm{\Delta u}=\bm{u}-\bm{u}_0$ the control law from Equation \ref{eq:u_control_law_indi} represents an incremental version of Equation \ref{eq:u_control_law}. Instead of computing the complete control input command $\bm{u}$, this control law results in computing the increment of the control input $\bm{\Delta u}$ and adding it to the previous value $\bm{u}_0$. As it is less dependent on the model of the system dynamics, it is able to increase the robustness of the system \cite{Sieberling}.

\subsection{Inner loop INDI for a Quadrotor} \label{subsec: indi_inner}

In this subsection, using theory from \ref{subsec:indi_general} the inner loop \acrshort{INDI} controller is derived to control angular acceleration of the Crazyflie quadrotor. The derivation of the inner loop, as well as the outer loop controllers is based on the \acrshort{INDI} controller architecture introduced by Smeur et al. \cite{Smeur1}, \cite{Smeur2}. Equation \ref{eq:ang_momentum} from section \ref{sec:eqs_motion} serves as a basis for this derivation. The desired variable to be controlled by the inner loop \acrshort{INDI} is the angular acceleration of the quadrotor in the body-fixed coordinate frame. As in the case of a real quadrotor control problem the value of the thrust can be seen as an output of the dynamic system, it makes sense to incorporate thrust as a control variable into the control law as well. Thus, the angular momentum Equation \ref{eq:ang_momentum} is augmented with the total thrust $T$ of all four rotors \cite{Smeur2}. As only the body-fixed frame is used in the following control law derivation, the subscripts $B$ are not used in this subsection. Furthermore, all external forces and moments apply to the center of gravity of the quadrotor, thus the superscripts $G$ are omitted. 

@@

@@

As already explained in the previous subsection, to derive the incremental control law the Taylor series expansion is perfromed on Equation \ref{eq:main_eom_indi_inner}

@@

By applying differentiation and rewriting some of the terms, the following equation is obtained:

@@

	\bm{G}_1 (\bm{\omega}) =

@@

			\omega_y & 0 & 0 \\
			0 & \omega_x & 0 \\

@@

With its arm length of approx. $4~\si{\mm}$, the Crazyflie 2.1 quadrotor developed by Swedish company Bitcraze belongs to the \acrfull{MAV} class (see Figure \ref{fig:crazyflie_drone}) \cite{petricca}. Its total weight is approx. $27~\si{g}$  with a capability to increase it up to $42~\si{g}$, which makes it possible to extend the default hardware with a pair of additional extension decks of various functionality. As it is an open source project, the main firmware, development tools and other software packages are freely available. This is one of the reasons why this small quadrotor gained a lot of popularity in different academia fields, being used for tasks reaching from the classical control theory all the way down to the obstacle avoidance problems \cite{bitcraze-research}.

The Crazyflie's main application \acrfull{MCU} is STM32F405. It runs the firmware with frequency of $1000~\si{\Hz}$ and communicates with a large number of peripheral devices such as Bluetooth and long range radio, power management system \acrshort{MCU}, \acrfull{IMU} etc.. The \acrfull{IMU} consists of the 3-axis accelerometer, the 3-axis gyroscope and the high precision pressure sensor.

The quadrotor used in this project was extended with a flow deck sensor. The flow deck itself is equipped with two sensors: the optical flow sensor which measures the visual motion and the \acrfull{ToF} laser-ranging sensor which measures the range to the next object along the body-fixed z-axis of the quadcopter. By employing these sensors, the flow deck can provide estimates of the current position and linear velocity of the aircraft.

\begin{figure}[H]
	\centering 
	\begin{tikzpicture}
		\node[inner sep=0pt] (crazyflie_drone) at (0,0) {
		\includegraphics[width=0.37\textwidth]{figs/crazyflie_drone.jpg}};
	\end{tikzpicture}
	% Figure description is centered aligned
	\captionsetup{justification=centering, singlelinecheck=off, font=bf, belowskip=-0.5cm}
	\caption[Crazyflie 2.1 nano-quadrotor]{Crazyflie 2.1 nano-quadrotor \cite{bitcraze}.}
	\label{fig:crazyflie_drone}
\end{figure}

@@

@@

This chapter describes the performance of the implemented outer loop. In section \ref{sec:disturbance_rejection} the disturbance rejection is experimentally analysed. Section \ref{sec:pos_controller_response} presents the general response of the position controller for different step inputs.
